---
output:
  github_document: default
  html_document: default
---

---
title: "hw_05"
author: "Barbara Klein"
output: github_document


## R Markdown

######## DIRECTIONS ###########
#Use ?nycflights13::flights to see the documentation of flights. (You can’t really show this in an RMarkdown document)
#Use the glimpse() function to get a general sense of the flights data frame. How many rows and columns does flights have?
#Use filter() to create a data frame of just the flights that departed 2 hours late or more. What fraction of flights departed at least 2 hours late?
#Use mutate() and paste0() to create a column that contains the combined airline and flight number (e.g., “UA5041”).
#Use group_by(), summarise(), and n() to make a data frame showing how many flights each aircraft (tail_num()) flew for each carrier throughout the whole data set.

################################# START OF HW-05 ######################################
```{r setup, echo=FALSE}
library(tidyverse)
library(dplyr)
library(nycflights13)
```

```{r glimpse, echo=TRUE}
?nycflights13::flights
glimpse(flights)

#There are 336,776 
#There are 19 Columns
```

```{r}

# Filter and create a data frame of just the flights that departed 2 hrs late or more (this means 2 hours AND more than that, so >= 2)
dep_delay2hrs <- filter(flights, dep_delay >= 2)
glimpse (dep_delay2hrs)

# Out of 336,776 flights, 120,382 of them departed two hours late or more (so ~35% of flights).
```


```{r}
# Use mutate and paste functions to make a whole new column with carrier and flight number combined.
flight_carrier_combo <- mutate(flights, 
                               carrier_flights_combo = paste0(carrier, flight))
glimpse(flight_carrier_combo)
# Now there is an extra column all the way at the bottom (20 instead of 19).
```
                                  
```{r}
# make a data frame showing how many flights each aircraft (tail_num()) flew for each carrier using the group_by(), summarise(), and n() functions. 
grouped_flights <- flights %>%
  group_by(carrier, tailnum) %>%
  summarise(grouped_flights = n()) 
glimpse(grouped_flights)  

#Now there are 4,067 rows, 3 columns, and 16 unique groups. 
#This is a test commit for Cameron          
```

