---
title: "HW10"
author: "Barbara Klein"
date: "11/11/2020"
output: html_document
---

## PART 1 - Linear models: 
#### Using the diamonds dataset, create a linear model of price as a function of the weight of the diamond(carat).

```{r, include=TRUE}
diamond_price <- lm (formula = price ~ carat, 
                     data = diamonds)

summary (diamond_price)
```

- According to that model, what price would you expect for a 1 - carat diamond? 
- ANS = The price for a 1 carat diamond would be $5,500.07 (7756.43 - 2256.36 = 5,500.07).


#### Now create a linear model of price as a function of carat and color. 
```{r}
diamond_color_price <- lm (price ~ carat + color, 
                           data = diamonds)

summary(diamond_color_price)
```

- What difference does it make if you convert color to a character vector before you make the model? (Note that this is probably what you want to do in general, unless you really know what you're doing.)
ANS = Color is being represented as a number (originally), in R. The diamonds$color summary show color as an ordered factor and R has automatically assigned it that way even though I didn't ask it to. So to convert it to a character vector means it isn't automatically ordered by R since we told R it's a character, it's just assigned a different value for color. So a color value of "E", would correspond to a price value, and so on. 
  
```{r}
diamond_color_chr_price <- lm (formula = price ~ carat + as.character(color), 
                               data = diamonds)

summary(diamond_color_chr_price)
```

  

- Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful. (Bonus: Should the data have been pre-processed before making these models?)
- ANS = No I don't think the data should've been pre-processed before making the model, because even as I have ggplot taking information from my manipulated 'diamonds' dataframe(s), the plots all appear the same. I'm not sure if this is an error on my part, or if it's just a caviat in R that I' don't'm unaware know of. 
  
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = lm)
```

    
```{r}
ggplot(diamond_color_price, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = lm)
```
  
```{r}
ggplot(diamond_color_chr_price, aes(x = carat, y = price)) +
  geom_point () +
  geom_smooth(method = lm)
```
  
  
# PART 2 - Objects and Methods:
#### Write a function that returns some kind of structured result, which you define as a new S3        class. Your class doesn’t need to be anything fancy, but it should be different from existing         classes.
```{r}
childhood_toys <- function(q, yoyo = "It's the thing on a string") {
  
  # check and make sure the input is a data frame; if not, run stop-message:
  if(!is.data.frame(q)) {
    stop("Initial argument is not a data frame. It needs to be a data frame")
  }
  
  if(!is.character(yoyo)){ 
    stop("yoyo must be a character")
  }
  
  class(q) <- c("funforhours_df", "data.frame")  #I create a class called "funforhours_df"
  attr(q, "yoyo") <- yoyo
  q
}
```

Now I can type the following down in the console:

- yoyo_fun <- childhood_toys(mtcars)

- class(yoyo_fun)

It returns: 

- [1] "funforhours_df" "data.frame"    


#### Write, print, and summarise methods for this class. These methods don’t need to be particularly useful, but they should be different from the default methods.
```{r}
summary.funforhours_df <- function(q) {
  yoyo <- attr(q, "yoyo")
  print(paste0("yoyo's are fun for many hours - ", yoyo))
  print(summary.data.frame(q))
  summary.data.frame(q)
}
#This function is taking the "yoyo" attribute, and printing 'yoyo's are fun for many hours' and then gives the data frame summary. 
```


Now I can type the following down in the console:

- "summary(yoyo_fun)"

And it gives us the summary [1] "yoyo's are fun for many hours - The thing on a string" along with     the mtcars dataframe. 

Here I use the sloop package to call the summary to the s3 dispatch function, on the function         called 'yoyo_fun'.

- "sloop::s3_dispatch(summary(yoyo_fun))

- => summary.funforhours_df

- *summary.data.frame

- *summary.default"

It successfully found the summary method for 'yoyo_fun'

##### It doesn't say to do this in the HW, but in following along with the video, and I want to validate my class, with a validator function.
```{r}
val_funforhours_df <- function(q) {
  if(is.null(attr(q, "yoyo"))) {
    stop("yoyo attribute is not there, or is missing")
  }
  
  # make sure it's a valid data frame. 
  if(!is.data.frame(q, "yoyo")) {
    stop("yoyo needs to be a data frame")
  }
}
```

