title: "hw_08"
author: "Barbara Klein"
output:
  github_markdown: default
  html_document: default
 

# **Tidyverse VS Base R**
### **Lists, data frames, and vectors**

* 1. Create a list: 

  + NAF = Name 
  + Age 
  + Favorite Food 
  
```{r}
NAF <- list ("name" = "Greg", "age" = "32", "FavFood" = "Apple Pie")
NAF $ name #Using '$' retrieves the content of the "name" element of the sublist within the NAF list.
NAF [[2]] #Retrieves the content of the list called "NAF", and even further to sublist #2, or in this case "age". Using two brackets are needed when accessing super specific elements of my list, or any list. If one were trying to access the second element of the list 'NAF', it could be accessed by either the name, or the number of the element.


NAF ["FavFood"] #gives us the sublist - "FavFood", not what's in the whole list of NAF.
```

* 2. Access mpg column in mtcars two separate ways

```{r}
#dplyr needed

data(mtcars) 
#Retrieving the mtcars data in tidyverse

mtcars %>% #showing the mtcars dataset table
  pull(mpg) 
#%>% is the pipe operator that takes the output of "mtcars" and makes it the input of 'pull mpg'. In other words you take the mtcars data, THEN (Then=%>%) you pull the mpg data within mtcars. 

mtcars %>%
  select(1) 

#I did the same as above but this time 'selected' the mpg column in the mtcars dataset. I knew it was the first column as shown by the mtcars dataset table above. 
```

* 3. In base R, we will now take the 'disp' column and divide it by the 'cyl' column, to create a new column called 'disp_per_cyl'. Here we will use the '[' to accomplish this transformation.

```{r}
mtcars[,"disp_per_cyl"] <- mtcars[, "disp"] / mtcars[, "cyl"]

#Another way to do it would be with the '$' approach: mtcars$disp_per_cyl <- mtcars$disp / mtcars$cyl
#A better/easier way to do it, outside of base R, would be to use the transform function: displacement_per_cylinder <- transform(disp_per_cyl = disp/cyl)

as_tibble(mtcars)
```


* 4. Create a single-column data frame (tibble) using the mutate function from tidyverse. 
```{r}
singlecol_dataframe <- mutate(mtcars, (disp_cyl = disp / cyl))
 singlecol_dataframe %>% select(-(1:11)) 
```
#### When using "select (-(1:11)), I am telling R to drop columns 1 through 11 to give me only column 12.
#### The difference here is I haven't created a new tibble from column vectors or individual vectors with, for example, the 'tibble()' command, but rather I've just mutated the current tibble/data frame I was already working in - mtcars. 

## **Factors**
* 1. Creating a numeric vector:

```{r}
#The vector must have all of the same type of data, they cannot mix or be different from one another. An example would be that all of these elements are integers (numbers), and are not a mix of integers and characters. 
num.vec <- c(3, 4, 1, 5)
```

* 2. To convert the newly created numeric vector to a factor we use the following:

```{r}
as.factor(num.vec)
```
#### The output mentioned in the console says there are four different "levels: 1 3 4 5" in chronological order. If the integer '1' were to be repeated in the first command "num.vec <- c(3, 4, 1, 5, 1)", we would see the same four levels in chronological order being listed, even though there are five total integers. 
* 3. To convert the newly created factor back to a numeric vector, we use the following:

```{r}
as.numeric(num.vec)
```
#### In this instance, we have taken away the 'level' portion of the data by calling the 'num.vec' a numeric vector again, rather than a factor. The factors identified above, are data objects that are categorized as levels, both in the form of integers or even strings. Data frames can comprise of multiple data types, which are in essence, vectors, which are then comprised of as factors. 

## **Exploratory data analysis with Figures**
```{r}
library(tidyverse)
library(nycflights13)
print(nycflights13::weather) 
#Since there are many data frames within the nycflights1, I have accessed the data frame specific to recorded weather characteristics. It prints a tibble with 15 columns and 26, 115 rows.


view(weather) 
#This command opens a new window in the viewer containing the dataset. This is what I will be using to select different columns, integers, or characters to re-arrange or manipulate in different ways. 


#Next I want to filter and focus on two months of 2013: December vs. June. 
Dec_Jun <- filter(weather, (month == 12 | month == 6))
print(Dec_Jun)


#Now I want to create a sublist of the days within Jun and Dec, and their respective wind direction, speed, and gust measurements. I will do this using the sublist I just filtered 'Dec_Jun' 


Dec_Jun_wind <- select(Dec_Jun, day, wind_speed, wind_gust)
#I see there are a lot of NA's in the viewer window of my selected weather data. I want to drop all NA values in this column since they give no information I want. 

Dec_Jun_wind %>% drop_na(wind_gust) 

?nycflights13::weather

#Next I want to plot (with ggplot2) wind speed and gusts on all days within the months of Jun and Dec, respectively. 
Dec_Jun_wind_plot <- ggplot(Dec_Jun_wind, aes(x = day, 
                         y = wind_speed, 
                         color = wind_gust)) + 
  geom_point(color = "darkred")
print(Dec_Jun_wind_plot)


#The 'plus' means I am adding "layers" to my ggplot
#This is a poorly made scatterplot, so I will make a tab and create another scatterplot. 
```



```{r}
ggplot(Dec_Jun_wind, aes(x = day, 
                         y = wind_speed)) +
  geom_point(mapping = aes(color = wind_gust)) +
  scale_color_distiller(type = "seq", 
                       palette = "Greens", 
                       guide = guide_legend(reverse = TRUE), 
                       direction = 1) +
  labs(title = "Wind Speed and Gust on a Given Day In Jun and Dec, 2013", 
       x = "Day", 
       y = "Wind Speed (mph)",
      color = "Wind Gust (mph)")
```

