---
title: "hw_09_part2"
author: "Barbara Klein"
date: "11/1/2020"
output: html_document
---

##### 1. Write a function to calculate variance of a numeric vector.
```{r}
#Here I will create my numeric vector: numeric_vec <- as.numeric(c(1, 3, 45, 23, 44, 22, 23, 12, 43, 32, 14))

functionforvariance <-function(numeric_vec) {
  var_vector <- var(numeric_vec)
  var_vector
}

#after the code chunk was run, the 'functions' environment showed up in my global environment, and the variance to my numeric vectors appeared as a result of the function.
```

##### 2. Write a function (both_na), in which two vectors of the same length and returns the number of positions that have an NA in both vectors.

```{r}
#Here I will create my two vectors of the same length: vec1 <- as.numeric(c(1, 2, NA, 4, 5, 6, NA, NA, 9)) . . .  vec2 <- as.numeric(c(2, 4, NA, 8, 10, 12, NA, 16, 18))

functionforbothNA <- function(vec1, vec2) {
  all_na <- sum(is.na(vec1) & is.na(vec2))
  all_na
}
functionforbothNA(
  c(1, 2, NA, 4, 5, 6, NA, NA, 9),
  c(2, 4, NA, 8, 10, 12, NA, 16, 18)
  )
```
The output R gives, explains that within my numeric vectors of the same length, two of the positions are the same in that they both have NA values in the same place. 

##### 3. Write a function to create a specific kind of plot, with useful labels, etc., that can take a data frame as an argument. 

```{r}
plotz <- function(poohbear){
ggplot(aes(x=cyl, y=mpg)) +
  geom_point() + 
  labs(x = "Cylinder", y = "Miles per Gallon (US)")
  plotz
}
```

##### 4. Write a function that takes one argument and does the following:
###### a) if the argument is a numveric vector, multiplies every number in the vector by 2. 
###### b) if the argument is a character vector, sorts the vector into alphabetical order.
###### c) if the argument is anything else, issues an informative warning and returns the argument unchanged.

```{r}
# Here is a character vector: tigger <- as.character("eeyore", "hunny", "roo", "piglet")
# The numeric vector will be the one used from the previous question #1 (above)

winnie_the_pooh <- function(owl) {
  runn <- if(is.numeric(owl)) {
    owl * 2}
  else if(is.character(owl)) {
    sort(owl)}
  else if(!is.numeric(owl)) if(!is.character(owl)) {
    warning("Warning: Non-numeric vector or non-character entry made. Argument must be numeric or character vector")}
  runn
}
winnie_the_pooh(TRUE)
```

##### 5. Write a function that tries to calculate the mean of a vector using mean(). If that is successful, it returns the mean of the vector. If mean() throws an error, it issues an informative message and returns the input intact. 

```{r}
# I will use the numeric vector created previously in question #1. 
mean_vec <- function(taco) {
  mv <- tryCatch({
    mean(taco)
  }, error = function(e)
    warning("Warning: Non-numeric vector entry made. Argument must be numeric")
  )
  mv
}
```

##### 6. Use system.time() (or, if you want to be more advanced about it, the bench or microbenchmark packages) to compare how long it takes to add two, 1-million-element random vectors and store the result in a third vector: 

```{r}
#install microbenchmark package: this package wraps the 'system.time' function and makes it straight forward as it compares more than one function. 
million_vec1 <- sample(1:100, 
                       1000000, 
                       replace = TRUE) 
million_vec2 <- sample(1:100, 
                       1000000, 
                       replace = TRUE)
```

###### -Vectors 'million_vec1' and 'million_vec2' are numeric with one million elements in each. 
###### -These vector objects in 'million_vec1' and 'million_vec2' can be repeated.
###### -Never put in NULL where length would be zero..the object would start empty and slowly fill up the vector with integers. This takes up memory and lots of time. Alternatively, you never want to 'grow' a vector. 

```{r}
# a.Using any vectorized operation (e.g., `+`).
millvec_systime <- function() {
  c(sum(sample(1:100,
               1000000,
               replace = TRUE)),
    sum(sample(1:100,
               1000000, 
               replace = TRUE)))
}
system.time(millvec_systime())
```

```{r}
# b.Using 'a' for loop with a pre-allocated vector to store the results.
store_millvec <- numeric()

system.time(store_millvec <- c(for(i in 1:1000000) 
                                    {i + i}
                               ))
```
```{r}
# c.Using 'a' for loop without pre-allocating a vector to store the results.
system.time(store_NA <- for(i in 1:1000000) 
                              {i + i}
                          )
```


